---
title: "COVID-10 sentiment analysis using Twitter"
author:   "by Mateo Cerviño"
date:     "`r Sys.Date()`"
output: html_notebook
---

First, installing packages
```{r}
install.packages("~/TextMining/Rstem_0.4-1.tar.gz", repos = NULL, type = "source")
install_url("http://cran.r-project.org/src/contrib/Archive/sentiment/sentiment_0.2.tar.gz")
require(sentiment)
ls("package:sentiment")
install.packages("tm")
install.packages("wordcloud")
install.packages("rtweet")
install.packages("igraph")
install.packages("graphTweets")
```

And load the libraries

```{r}
# library(twitteR)
library(rtweet)
library(wordcloud)
library(RColorBrewer)
library(plyr)
library(ggplot2)
library(sentiment)
library(ggplot2)
library(gridExtra)
library(wordcloud)
library(igraph)
library(graphTweets)
```


First, access to Twitter with the token created using the developer account info

```{r}

create_token(
  app = twitter_app,
  consumer_key = api_key,
  consumer_secret = api_secret,
  access_token = access_token,
  acces_secret = access_token_secret)
```

Request for 18000 tweets

```{r}
covid_media_tweets <- search_tweets(q = '-filter:replies filter:verified -filter:media covid OR pandemic OR quarantine OR pcr OR "covid testing"', type = "recent", n = 8000, include_rts = FALSE)
```

Take a look at the returned text without personal accounts or html links

```{r}
covid_media_tweets_txt <- covid_media_tweets$text
covid_media_tweets_txt <- gsub("@\\w+", "",covid_media_tweets_txt) #without personal accounts
covid_media_tweets_txt <- gsub("\\bhttp[a-zA-z0-9]*\\b", "",covid_media_tweets_txt) #without html links
covid_media_tweets_txt <- gsub("[[:punct:]]", "", covid_media_tweets_txt) #without score
covid_media_tweets_txt <- gsub("[^a-zA-Z0-9 ]", "", covid_media_tweets_txt) #without non alphanumeric characters
covid_media_tweets_txt <-  gsub("\\btco[a-zA-Z0-9]*\\b", "", covid_media_tweets_txt) #without tco
covid_media_tweets_txt <- covid_media_tweets_txt[!is.na(covid_media_tweets_txt)] # without NAs
covid_media_tweets_txt <- iconv(covid_media_tweets_txt, 'UTF-8', 'ASCII') # without emojis
covid_media_tweets_txt <- tolower(covid_media_tweets_txt) # all in lower case
covid_media_tweets_txt <- gsub("[ \t]{2,}", "", covid_media_tweets_txt) # spaces and tabs
covid_media_tweets_txt <- gsub("^\\s+|\\s+$", "", covid_media_tweets_txt)
```

Look at the final text

```{r}
head(covid_media_tweets_txt, 10)
```


### Emotion classification

Starting classifying emotions with *classify_emotion* using Bayes algorithm

```{r}
covid_media_tweets_emo <- classify_emotion(covid_media_tweets_txt, algorithm="bayes", prior=1.0)
```

The function returns 7 columns: anger, disgust, fear, joy, sadness, surprise and best_fit for every file in the document

```{r}
head(covid_media_tweets_emo)
```

Save the (BEST_FIT) 

```{r}
# Save object emotion
emotion <- covid_media_tweets_emo[, 7]

table(emotion, useNA = 'ifany')
```

To facilitate the classification, change NAs for unknown

```{r}
emotion[is.na(emotion)] <- "unknown"
table(emotion, useNA = 'ifany')
```

In a graphic 

```{r, fig.align="center"}
pie <- ggplot(as.data.frame(covid_media_tweets_emo), aes(x = factor(1), fill = factor(BEST_FIT))) +
 geom_bar(width = 1)
pie + coord_polar(theta = "y") + labs(title = 'Semtiment COVID')
```

```{r, fig.align="center"}
g <- ggplot(as.data.frame(covid_media_tweets_emo), aes(x = BEST_FIT)) +
 geom_bar() + labs(title = 'Semtiment COVID') + geom_bar(aes(y=..count.., fill=emotion)) +
      scale_fill_brewer(palette="Dark2")
g 
```

Now we take a look at the polarity. We classify the text in four categories

* POS:      Positive
* NEG:      Negative
* POS/NEG:  Undefined
* BEST_FIT: Common

```{r}
covid_media_tweets_pol <- classify_polarity(covid_media_tweets_txt, algorithm="bayes")
head(covid_media_tweets_pol, 3)
```

Create an object with the result, polarity

```{r}
polarity <- covid_media_tweets_pol[, 4]
head(polarity)
```

```{r}
table(polarity, useNA = 'ifany')
```

All the info in a dataframe

```{r}
sentiment_dataframe <- data.frame(text     = covid_media_tweets_txt, 
                                  emotion  = emotion, 
                                  polarity = polarity, stringsAsFactors=FALSE)
head(sentiment_dataframe)
```

Reorder it

```{r}
sentiment_dataframe <- within(sentiment_dataframe, emotion <- factor(emotion, levels=names(sort(table(emotion), decreasing=TRUE))))
head(sentiment_dataframe)
```

### Graphics

Number of tweets classified by categories

```{r, fig.align="center"}
ggplot(sentiment_dataframe, aes(x = emotion)) + geom_bar(aes(y = ..count.., fill = emotion)) +
      scale_fill_brewer(palette = "Dark2") +
      ggtitle("Sentiment analysis COVID on Twitter") +
      theme(legend.position="right") + ylab("Number of Tweets") + xlab("Emotion types")
```

Polarity of tweets

```{r, fig.align="center"}
ggplot(sentiment_dataframe, aes(x = polarity)) +
      geom_bar(aes(y = ..count.., fill = polarity)) +
      scale_fill_brewer(palette = "RdGy") +
      ggtitle("Sentiment analysis COVID on Twitter") +
      theme(legend.position="right") + ylab("Number of Tweets") + xlab("Polarity types")
```

Split the words according to the emotions

```{r}
covid_media_tweets_emos     <- levels(factor(sentiment_dataframe$emotion))
n_covid_media_tweets_emos  <- length(covid_media_tweets_emos)
covid_media_tweets_emo_docs <- rep("", n_covid_media_tweets_emos)
for (i in 1:n_covid_media_tweets_emos)
{
      tmp <- covid_media_tweets_txt[emotion == covid_media_tweets_emos[i]]
      covid_media_tweets_emo_docs[i] <- paste(tmp, collapse=" ")
}
```

```{r}
head(covid_media_tweets_emos)
```

Creating a corpus

```{r}
covid_media_tweets_corpus <- Corpus(VectorSource(covid_media_tweets_txt))
inspect(covid_media_tweets_corpus[1:10])
```

Clean it with *tm_map*

```{r}
# lower case
corpus_clean <- tm_map(covid_media_tweets_corpus, tolower)

# Without numbers
corpus_clean <- tm_map(corpus_clean, removeNumbers)

# Without score
corpus_clean <- tm_map(corpus_clean, removePunctuation)

# Without stopwords
corpus_clean <- tm_map(corpus_clean, removeWords, stopwords())

# Without blank spaces
corpus_clean <- tm_map(corpus_clean, stripWhitespace)
```

Un tdm (Term Document Matrix)

```{r}
covid_media_tweets_tdm           <- TermDocumentMatrix(covid_media_tweets_corpus, control = list(stopwords = TRUE))
covid_media_tweets_tdm           <- as.matrix(covid_media_tweets_tdm)
```

A dtm (Document-Term Matrix)

```{r}
covid_media_tweets_dtm <- DocumentTermMatrix(covid_media_tweets_corpus, control = list(minWordLength = 1, 
                                                       stopwords = TRUE))
inspect(covid_media_tweets_dtm)
```

Stem

```{r}
covid_media_tweets_corpus_stem <- tm_map(covid_media_tweets_corpus, stemDocument)
covid_media_tweets_corpus_stem <- tm_map(covid_media_tweets_corpus_stem, stemCompletion, dictionary = covid_media_tweets_corpus)
inspect(covid_media_tweets_corpus_stem[1:5])
```

Frequent terms

```{r}
head(findFreqTerms(covid_media_tweets_dtm, lowfreq=10), 20)
```

Words associated to word blood

```{r}
findAssocs(covid_media_tweets_dtm, 'blood', 0.30)
```


### Word cloud

```{r, fig.align="center"}
wordcloud(corpus_clean, min.freq = 20, random.order = FALSE)
```










